---
title: "Assignment week 05"
author: "Sara Leka"
date: "2025-03-26"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(AnnotationHub)
  library(ensembldb)
  library(GenomicRanges)
  library(epiwraps)
  library(rtracklayer)
  library(ggplot2)
})
```

## Using data from from mouse embryonic stem cells (mESC)

### Histone modification mark H3K4me3
```{r, eval=FALSE}
dir.create("mesc", showWarnings=FALSE)
download.file("https://www.encodeproject.org/files/ENCFF974BMC/@@download/ENCFF974BMC.bed.gz",dest="H3K4me3.bed.gz")
```

```{r}
H3K4me3 <- rtracklayer::import("H3K4me3.bed.gz", format="narrowPeak")
head(H3K4me3)
```

### Histone modification mark H3K27me3
```{r,eval=FALSE}
download.file("https://www.encodeproject.org/files/ENCFF008XKX/@@download/ENCFF008XKX.bed.gz",dest="H3K27me3.bed.gz")
```
```{r}
H3K27me3 <- rtracklayer::import("H3K27me3.bed.gz", format="narrowPeak")
head(H3K27me3)
```

## 1. Bivalent domains identification

```{r}
overlapping_indices <- findOverlaps(H3K27me3, H3K4me3, type = "any")
bivalent_regions <- pintersect(H3K27me3[from(overlapping_indices)], H3K4me3[to(overlapping_indices)])
cat("Number of bivalent regions:", length(bivalent_regions), "\n")
```

## 2. Behavior upon differentiation
### Using B cell data

```{r, eval=FALSE}
download.file("https://www.encodeproject.org/files/ENCFF496DOC/@@download/ENCFF496DOC.bed.gz", "H3K4me3_Bcell.bed.gz")
download.file("https://www.encodeproject.org/files/ENCFF931RRP/@@download/ENCFF931RRP.bed.gz", "H3K27me3_Bcell.bed.gz")
```
```{r}
B_H3K4me3 <- import("H3K4me3_Bcell.bed.gz", format = "narrowPeak")
B_H3K27me3 <- import("H3K27me3_Bcell.bed.gz", format = "narrowPeak")
```
### Bivalent regions that only overlap with H3K4me3 marks in B cells
```{r}
H3K4me3_only <- findOverlaps(bivalent_regions, B_H3K4me3, type = "any")
length(H3K4me3_only)
```

### Bivalent regions that only overlap with H3K27me3 marks in B cells
```{r}
H3K27me3_only <- findOverlaps(bivalent_regions, B_H3K27me3, type = "any")
length(H3K27me3_only)
```

### Bivalent regions that overlap with both marks in B cells
```{r}
overlap_B <- findOverlaps(B_H3K4me3, B_H3K27me3, type = "any")
both_domains_B <- pintersect(B_H3K4me3[from(overlap_B)], B_H3K27me3[to(overlap_B)])

overlap_both <- findOverlaps(bivalent_regions, both_domains_B, type = "any")
length(overlap_both)
```
```{r}
cat("Bivalent regions that only overlap with H3K4me3 marks in B cells: ", length(H3K4me3_only),
    "\nBivalent regions that only overlap with H3K27me3 marks in B cells: ", length(H3K27me3_only),
    "\nBivalent regions that overlapping with the combination of the two marks in B cells: ",  length(overlap_both))
```


