---
title: "Assignment Week 06"
author: "Sara Leka"
date: "2025-04-02"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(AnnotationHub)
  library(GenomicRanges)
  library(ggplot2)
  library(motifmatchr) # for scanning sequences for matches of given motifs
  library(Biostrings) # for handling sequences
  library(MotifDb) # database of motifs
  library(TFBSTools) # for handling some motif formats
  library(universalmotif) # for converting motifs to various formats
  library(PWMEnrich) # for R-based motif enrichment analysis
  library(BSgenome.Mmusculus.UCSC.mm10)
})
```


## I chose CEBPB TF ChIP-seq data from Mus Musculus myocytes (assembly mm10)


### Downloading the peaks for the factor (only chromosome 1 for faster computation)
```{r}
download.file("https://www.encodeproject.org/files/ENCFF980ZHP/@@download/ENCFF980ZHP.bed.gz", "Mus_CEBPB_myocytes_ENCFF980ZHP.bed.gz")
peaks <- rtracklayer::import("Mus_CEBPB_myocytes_ENCFF980ZHP.bed.gz", format="NarrowPeak")
seqlevelsStyle(peaks) <- "UCSC"
peaks_chr1 <- peaks[seqnames(peaks)=="chr1"]
```

### Getting the genome sequence (I actually used BSgenome because I had problems with the genome.fa file)
```{r, eval=FALSE}
#ah <- AnnotationHub()
#AnnotationHub::query(ah, c("mm10", "2bit"))

#AH14005

#genome <- ah[["AH14005"]]
#genome_seqs <- import(genome)
```

### Identify the instances of the factor’s motif

```{r}
motifs <- query(MotifDb, "CEBPB")
names(motifs)

#Mmusculus-HOCOMOCOv10-CEBPB_MOUSE.H10MO.A

motif <- motifs[["Mmusculus-HOCOMOCOv10-CEBPB_MOUSE.H10MO.A"]]
view_motifs(motifs[1:2])
```

### Of all the peaks, what proportion contains a motif for the factor?


```{r}
motif2 <- convert_motifs(motif, class="TFBSTools-PWMatrix")
moi <- motifmatchr::matchMotifs(motif2, subject=peaks_chr1, genome=BSgenome.Mmusculus.UCSC.mm10,out="positions")
moi <- moi[[1]]
head(moi)
hist(moi$score)
```

```{r}
table(overlapsAny(peaks_chr1, moi))
```

```{r}
peak_perc <- sum(overlapsAny(peaks_chr1, moi))/length(peaks_chr1)*100
cat("\nOf the ",length(peaks_chr1)," peaks in chromosome 1, ",sum(overlapsAny(peaks_chr1, moi))," (",peak_perc,"% ) contain a motif for CEBPB")
```

### Of all instances of that motif in the genome (or in one chromosome), what proportion is bound by the factor (i.e. has a peak)?

```{r}
motif_perc <- sum(overlapsAny(peaks_chr1, moi))/length(moi)*100
cat("\nOf the ",length(moi)," motif instances in chromosome 1, ",sum(overlapsAny(peaks_chr1, moi))," (",motif_perc,"% ) have a peak")
```